---
title: "Microbial Sunburst Plots"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Microbial Sunburst Plots}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Aim

To build a sunburst chart that represents microbial composition of some microbiome sample. Our input will be a vector of ncbi taxonomy IDs (taxid).

A taxids abundance = the frequency in input vector



# Libraries

We'll need two R packages [sunburst](https://github.com/selkamand/sunburst) and [taxizedbextra](https://github.com/selkamand/taxizedbextra)

```{r setup}
# remotes::install_github("selkamand/sunburst")
library(sunburst) 

# remotes::install_github("selkamand/taxizedbextra")
library(taxizedbextra) 
```

# Download ncbi taxonomy database

First, lets use \`db_download_ncbi()\` from `taxizedb` (exposed by `taxizedbextra`) to download the ncbi taxonomy database locally to help us build sunburst plots from taxonomy ids at blistering speed. Its just over 2 gigs so might take a while, but it'll be worth it down the road.

```{r, eval=FALSE}
# Download ncbi taxonomy database. 
db_download_ncbi(overwrite = TRUE)
```

On my macbook it saves the db to `~/Library/Caches/R/taxizedb`. You can check where its downloaded for you by running `locate_taxonomy_cache()`

```{r}
# Where is my taxonomy database downloaded to?
locate_taxonomy_cache()
```

# Get data for sunburst plot

We need to get data in the required format (numeric vector of ncbi taxids)
You can use taxid = -1 for 'unclassified sequences'

```{r}
# Here we're simulating our data
taxids = c(rep(561, times = 10), rep(1639, times = 20), rep(529731, times = 10))

taxids
```


# Create sunburst plot
```{r fig.height=7, fig.width=7}
# generate sunburst plot
microbial_sunburst(
  taxids = taxids, 
  ranks_to_include = c("species", "genus", "family")
  )
```

run `?microbial_sunburst()` to learn how to customise this plot


# Acknowledgements

The engines driving the functionality of this package are the `taxizedb` and `plotly` packages. A big thanks to all involved in the creation and maintenance of these packages.
